---
title: "Mulitvariate"
author: "Wen Yang"
date: "February 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Multi-variate Plot & Analysis

Observations:

```{r}
ggplot(dta, aes(y=citric.acid, x=fixed.acidity, colour=level)) +
  geom_point(alpha=0.3, position = position_jitter()) +
  stat_smooth(method="lm", se=FALSE)
```

* at given fixed.acidity level, the higher citric.acid, the higher wine quality level. citirc.acid is indeed a import facotrs contributing to wine quality
* 

Let's move to another strongly positive correlated group fixed.acidity and density:

```{r}
ggplot(dta, aes(y=density, x=fixed.acidity, colour=level)) +
  geom_point(alpha=0.3, position = position_jitter()) +
  stat_smooth(method="lm", se=FALSE)
```

Observations:

* Given same density, the higher the fixed.acidity, the higher wine quality

Check the Strong correlated group in terms of level:

```{r}
# fixed.acidity and citric.acid has positive strong correlation

ggplot(dta, aes(x=citric.acid,y=alcohol,  colour=level)) +
  geom_point(alpha=0.3, position = position_jitter()) +
  stat_smooth(method="lm", se=FALSE)
```





```{r}
ggplot(dta, aes(y=citric.acid, x=log10(residual.sugar), colour=level)) +
  geom_point(alpha=0.3, position = position_jitter()) +
  stat_smooth(method="lm", se=FALSE)
```


```{r}
ggplot(dta, aes(x=log10(residual.sugar), y= citric.acid, colour=level)) +
  facet_wrap(~ level) +
  geom_point(alpha=0.3, position="jitter") +
  geom_boxplot(alpha=0, colour="black")
```